A 3d curvilinear skeletonization algorithm with application to path tracing

### 2012-09-11 skeleton_and_pathtracing.tex
Corrections de michel, intro en cours

\def\birth#1#2{Birth_{#2}(#1)}
\def\death#1#2{Death_{#2}(#1)}
\def\lifespan#1#2{Lifespan_{#2}(#1)}

\def\V2C{\Psi}
\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Nset{\mathbb{N}}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\cont#1{\check{#1}}
\def\contstrict#1{\check{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\detach#1#2{#2 \oslash #1}
\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}Border\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Noise in photorealistic computer generated images. Need to accelerate convergence. Add knowledge of the geometry and topology of the scene thanks to discrete geometry. Use the skeleton. Need of a robust curvilinear skeleton that represent topology and approximate geometry. Use the cubical complex framework.

\section{The cubical complex framework}

\subsection{Basic definitions}

@@

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.

Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ is \myem{a face of $f$}, and any element of $\cellstrict{f}$ is \myem{a proper face of $f$}. We set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and $\contstrict{f} = \cont{f} \setminus \{f\}$. It is plain that $g \in \cell{f}$ iff $f \in \cont{g}$.

A set $X$ of faces in $\allfaces{n}$ is a \myem{cell}, or \myem{$m$-cell}, if there exists an $m$-face $f \in X$ such that $X=\cell{f}$. The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$. The set $\Compl{X}$ is $\allfaces{n} \setminus X$.

@@

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. We denote by $\princ{X}$ the set composed of all facets of $X$. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.

\subsection{From binary images to cubical complex}
\label{sec::binary_to_cubic}

Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$). We give now a simple way to transpose such image to the cubical complex framework.

To do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$. More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\V2C(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\V2C$ to sets: $\V2C(S)=\{\V2C(x) | x \in S\}$. Given a set $S \subset \Zset^n$, we associate to it the cubical complex $\closure{\V2C(S)}$.

\subsection{Thinning: the collapse operation}
The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they are free:

\begin{definition}
Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $(\contstrict{g} \inter X) = \{f\}$. 
\end{definition}

@@

\subsection{Removing free pairs in parallel}
In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we give now. First, we need to define the \myem{direction} and the \myem{orientation} of a free face.

Let  $f \in \allfaces{n}$, the \myem{center of $f$} is the center of mass of the points in $f$, that is, $c_f= \frac{1}{\Card{f}}\sum_{a \in f}{a}$. The center of $f$ is an element of $[\Zhalf]^n$, where $\Zhalf$ denotes the set of half integers.
Let $\complex{n}{X}$, let $(f,g)$ be a free pair for $X$, and let $c_f$ and $c_g$ be the respective centers of the faces $f$ and $g$. We denote by $V(f,g)$ the vector $(c_f -c_g)$ of $[\Zhalf]^n$.

We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear. The number $\type{f,g}$ is called the \myem{direction} of the free pair $(f,g)$. Let $(f,g)$ be a free pair, the vector $V(f,g)$ has only one non-null coordinate: the pair $(f,g)$ has a \myem{positive orientation}, and we write $\orient{f,g}=1$, if the non-null coordinate of $V(f,g)$ is positive; otherwise $(f,g)$ has a \myem{negative orientation}, and we write $\orient{f,g}=0$.
%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.

Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.

@@

\begin{proof}
If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$. Also, $(f,g)$ is not free for $X \setminus \{k,\ell\}$. If $f\neq k$, then we have $g \neq \ell$, $\contstrict{g} \inter X = \{f\}$ ($g$ is free for $X$) and $\contstrict{\ell} \inter X = \{k\}$ ($\ell$ is free for $X$). Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$ and $(k,\ell)$ is a free pair for $Y$. $\finproof$
\end{proof}

@@

\subsection{A directional parallel thinning algorithm}

Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer} and avoid having unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex. 
We define $\border{X}$ as the set all faces belonging to a free pair for $X$. We now introduce Alg.~\ref{algo::pardircollapse_v0}, a directional parallel thinning algorithm.

On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}). Thanks to corollary \ref{cor:parallel_removal_v2}, we remove faces with same direction and orientation in parallel (l.~\ref{pc:08}).

@@

Different definitions of orientation and direction can be given, each corresponding to a different order of free faces removal in the complex and leading to different results.
Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces in the complex). Indeed, checking if a face is free or not may be easily done in constant time. Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $\cellstrict{f}$ in order to find new free faces.

@@

In the following, we give a new method requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory visual properties. This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0}, some faces from the original object during the thinning process.

\subsection{The lifespan of a face}
In the following, we define additional functions in the cubical complex, which are essential for the filtering step of the skeleton. The first one we present is the \myem{death date} of a face.

@@

The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$. We now define the \myem{birth date} of a face:

@@

The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of the largest maximal $1$-ball contained in $S$ and containing $x$. If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$. The map $\DD1{S}$ is known as the opening function of $S$ (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.

@@

In order to extend all these previous maps defined in $\Zset^n$ to the cubical complex framework, let us introduce the map $\V2C^{-1}$, inverse of the bijective map $\V2C : \Zset^{n} \rightarrow \subfaces{n}{n}$ defined in Sec.~\ref{sec::binary_to_cubic}. It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$. We indifferently use $\V2C^{-1}$ as a map from $\subfaces{n}{n}$ to $\Zset^n$, and as a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$.

@@

\bigskip
\begin{math}
\cplDist1{Y}(f) = \left\{
\begin{array}{ll}
\Dist1{S}(\V2C^{-1}(f)) & \text{if $f$ is an $n$-face}\\
\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \cplDist1{Y}(g) & \text{else}
\end{array}
\right.
\end{math}

@@

We can identify surfacic areas of a complex as zones where squares have a high lifespan. Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to appear in these zones. It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge. This leads us to proposing Alg.~\ref{algo::3d_curvskel}.

>>>>

\def\birth#1#2{\mbox{\it{}Birth\/}_{#2}(#1)}
\def\death#1#2{\mbox{\it{}Death\/}_{#2}(#1)}
\def\lifespan#1#2{\mbox{\it{}Lifespan\/}_{#2}(#1)}

\def\V2C{\Psi}
\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Nset{\mathbb{N}}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}Border\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation. The photorealistic aspect is achieved by simulating diffusion of light: elements of the scene which are illuminated by a light source can cast this light to other elements which are not directly lighted by any light source, thus giving a natural effect of light diffusion. To do so, a path tracing algorithm sends rays (of light) in the scene

The algorithm sends rays of light simulates global illumination effect: parts of the scene which are not directly illuminated by a light source the idea behind the algorithm is very elegant: for each pixel of the image, it throw rays through the pixel and let it bounce over in the scene (like a real ray of light would do). At each bouncethat will bounce on the scene to accumulate lighting exchanges. In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights. Our idea is to use a curvilinear skeleton of the void of the scene to guide the rays in the correct direction. We start by defining some useful concepts related to path tracing and we present our skeleton based path tracing algorithm.

Noise in photorealistic computer generated images. Need to accelerate convergence. Add knowledge of the geometry and topology of the scene thanks to discrete geometry. Use the skeleton. Need of a robust curvilinear skeleton that represent topology and approximate geometry. Use the cubical complex framework.

\section{The cubical complex framework}
\label{sec::binary_to_cubic}

@@

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ (resp. $\cellstrict{f}$) is \myem{a face of $f$} (resp. \myem{a proper face of $f$}.
The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$. The set $\Compl{X}$ is $\allfaces{n} \setminus X$.

@@

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.

Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$). To transpose such image to the cubical complex framework, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$. Let $x=(x_1,...,x_n) \in S$, we define the n-face $\V2C(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\V2C$ to sets: $\V2C(S)=\{\V2C(x) | x \in S\}$. Given a set $S \subset \Zset^n$, we associate to it the cubical complex $\closure{\V2C(S)}$.

The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they are free:

\begin{definition}
Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $f$ is the only face of $X$ which strictly contains $g$. 
\end{definition}

@@

In the cubical complex framework, parallel removal of free pairs can be easily achieved when following simple rules that we give now. First, we need to define the \myem{direction} and the \myem{orientation} of a free face.
Let $(f,g)$ be a free pair for $\complex{n}{X}$ : we have $\mydim{g} = \mydim{f}-1$, and it can be seen that $g = f \inter f'$, where $f'$ is the translate of $f$ by one of the $2n$ vectors of $\Zset^n$ which have all their coordinates equal to 0 except one, which is either equal to +1 or -1. Let $v$ be this vector, and $c$ its non-null coordinate. We define $\type{f,g}$, called the \myem{direction} of the free pair $(f,g)$, as the index of $c$ in $v$. The \myem{orientation} of the free pair $(f,g)$ is defined as $\orient{f,g}=1$ if $c=1$, and $\orient{f,g}=0$ else.

%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but different orientations.

Now, we give a property of collapse (previously proven in \cite{ChauCou2009}) which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.

@@

%\begin{proof}
%If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$. Also, $(f,g)$ is not free for $X \setminus \{k,\ell\}$. If $f\neq k$, then we have $g \neq \ell$, $\contstrict{g} \inter X = \{f\}$ ($g$ is free for $X$) and $\contstrict{\ell} \inter X = \{k\}$ ($\ell$ is free for $X$). Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$ and $(k,\ell)$ is a free pair for $Y$. $\finproof$
%\end{proof}

@@

Intuitively, we want our thinning algorithm to remove free faces of a complex \quotes{layer by layer} and avoid having unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex. 
We define $\border{X}$ as the set all faces belonging to a free pair for $X$. We now introduce Alg.~\ref{algo::pardircollapse_v0}, a directional parallel thinning algorithm.

On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}). Thanks to corollary \ref{cor:parallel_removal_v2}, we can remove faces with same direction and orientation in parallel (l.~\ref{pc:08}), while guaranteeing topology preservation.

@@

Different definitions of orientation and direction can be given, each corresponding to a different order of free faces removal in the complex and leading to different results.
Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces in the complex). Indeed, checking if a face is free or not may be easily done in constant time. Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces contained in $f$ in order to find new free faces.

@@

In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties. This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0}, some faces from the original object during the thinning process.

\subsection{The lifespan of a face}
In the following, we define additional functions in the cubical complex, related to the thinning process (Alg.~\ref{algo::pardircollapse_v0}), which are essential for the filtering step of the skeleton. The first one we present is the \myem{death date} of a face.

@@

Intuitively, the death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$. We now define the \myem{birth date} of a face:

@@

The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of the largest maximal $1$-ball contained in $S$ and containing $x$. If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$. The map $\DD1{S}$ is known as the opening function of $S$ based on the $1$-distance (also called the granulometry function) \cite{Matheron67}: it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.

@@

In order to extend all these previous maps defined in $\Zset^n$ to the cubical complex framework, we use the map $\V2C^{-1}$, inverse of the bijective map $\V2C : \Zset^{n} \rightarrow \subfaces{n}{n}$ defined in Sec.~\ref{sec::binary_to_cubic}. It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$. This map induces a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$, that we also denote by $\V2C^{-1}$.

@@

\bigskip
\begin{math}
\cplDist1{Y}(f) = \left\{
\begin{array}{ll}
\Dist1{S}(\V2C^{-1}(f)) & \text{if $f$ is an $n$-face}\\
\displaystyle \max_{f \in \cellstrict{g} \inter \subfaces{n}{n}} \cplDist1{Y}(g) & \text{otherwise}
\end{array}
\right.
\end{math}

@@

We can identify surfacic areas of a complex as zones where squares have a high lifespan. Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to be preserved in these zones. It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge. This leads us to proposing Alg.~\ref{algo::3d_curvskel}.
